---
title: "SDS315 HW 10"
author: "Juhi Malwade - jm97555"
date: "2024-04-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=7,fig.align = "left", warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

[Github Link](https://github.com/juhimalwade/SDS315.git)

## Problem 1: Redlining

### Question:
Is there an association between the number of FAIR policies and the racial/ethnic composition of a ZIP code, adjusting for the fire, age, and income variables?

### Approach:
To answer this question, I fit a multiple linear regression model that predicts the number of fair policies using minority, fire, age, and income as variables. It had the form: policies = \(\beta_0\) + \(\beta_1\) x minority + \(\beta_2\) x fire + \(\beta_3\) x age + \(\beta_3\) x income. The coefficient for minority, \(\beta_1\), represents the change in number of policies for a one unit increase in minority percentage holding the other variables constant. Thus, I will use this coefficient and a 95% confidence interval to answer the question above.

```{r,echo=FALSE,message=FALSE}
library(knitr)
redline <- read.csv("redlining.csv")

fit_redline = lm(policies ~ minority + fire + age + income, data=redline)
#fit_redline
kable(confint(fit_redline, level=0.95),caption = "95% Confidence Interval")
```
### Results:
The model's coefficient for minority, \(\beta_1\), was 0.008359. To account for statistical uncertainty, I determined the 95% confidence interval for \(\beta_1\) to be between 0.00254 and 0.01417. Because it did not include zero, the positive relationship between racial/ethnic composition and number of FAIR policies is statistically significant to the 0.05 level.

### Conclusion:
In conclusion, there is a positive association between the number of FAIR policies and racial/ethnic composition of a zip code, adjusting for the number of fires, ages of homes, and income. This indicates there is evidence for racial discrimination in the private home insurance market. 

<br>

## Problem 2: Grocery Store Prices

### Part A

```{r,echo=FALSE,message=FALSE}
library(ggplot2)
library(dplyr)

groceries <- read.csv("groceries.csv")

groceries_1 <- groceries %>% 
  group_by(Store) %>% 
  summarize(Avg_Price = mean(Price))

ggplot(groceries_1) + geom_col(aes(x=reorder(Store,Avg_Price),y=Avg_Price)) + coord_flip() + labs(title = "Average Price of Grocery Products", x = "Store", y = "Average Price ($)")


```

The bar graph above displays the average price of grocery products for 14 stores across Texas. According to the graph, Whole Foods and Wheatsville Co-Op have the most expensive groceries while Walmart and Fiesta have the cheapest groceries.

### Part B

```{r,echo=FALSE,message=FALSE}
groceries_2  = groceries %>% 
  group_by(Product) %>% 
  summarize(num_stores = n())

ggplot(groceries_2) + geom_col(aes(x=reorder(Product,num_stores),y=num_stores)) + coord_flip() + theme(axis.text.y = element_text(size = 5)) + labs(title = "Number of Stores selling each Product", y = "Number of Stores",x="Product")

```
The plot shows the number of stores selling each product. Horizon 2% Milk Carton and a carton of eggs is in all 16 stores. Lucky Charms, Frosted Flakes, El Milagros Tortilla Chips, and Cinnamon Toast Crunch are only in 4 stores.

### Part C

```{r, echo = FALSE,message = FALSE}
fit_groceries = lm(Price ~ Product + Type, data=groceries)

kable(confint(fit_groceries, level=0.95))
```

Compared with ordinary grocery stores (like Albertsons, HEB, or Krogers), convenience stores charge somewhere between 0.41 and 0.92 dollars more for the same product.

### Part D

```{r,echo=FALSE,message=FALSE}
fit_groceries_2 = lm(Price ~ Product + Store, data=groceries)
#confint(fit_groceries_2, level=0.95)
fit_groceries_2

```
Walmart and Kroger Fresh Fare seem to charge the lowest prices when comparing the same product. Whole Foods and Wheatsville Food Co-Op seem to charge the highest prices when comparing the same product.

### Part E

Central Market charges slightly more than HEB for the same product. However, the difference in price is fairly small, as Central Market had a coefficient of -0.57339 and HEB had a coefficient of -0.64596. This amounts to around a 7 cent difference in price. Other stores had significantly higher price differences. For example, the difference between Whole Foods (the most expensive store) and Walmart (the least expensive store) was around 30 cents. 

### Part F

```{r,echo=FALSE,message=FALSE}
groceries_3 <- groceries %>%
  mutate(Income10K = Income/10000)

fit_groceries_3 = lm(Price ~ Product + Income10K, data=groceries_3)
fit_groceries_3

```
The Income10K coefficient is -0.01409. Because the sign is negative, it shows that consumers in poorer ZIP codes seem to pay more for the same product, on average. This is because a negative coefficient in the model indicates that as the mean income of the zipcode increases, price seems to go down. Thus, richer ZIP codes are paying less for the same product.

```{r,echo=FALSE,message=FALSE}
library(effectsize)
kable(standardize_parameters(fit_groceries_3))

```

A one-standard deviation increase in the income of a ZIP code seems to be associated with a -0.03 standard-deviation change in the price that consumers in that ZIP code expect to pay for the same product.
